\subsection{MEPS Data}\label{sec:MEPS}

The MEPS is a short panel survey focusing on medical events and expenses incurred through them.  Respondents are interviewed every six months over a two year span, for a total of five waves; just under 90\% of individuals in the MEPS are interviewed and report their SRHS in all five waves.  A new MEPS panel begins every year, so two overlapping panels are interviewed in each data wave.  Households heads of any age can be selected for the MEPS, and the SRHS question is asked for all members of a sample household.  The dataset constructed for this paper uses Panels 1 through 20, spanning 1996 to 2016.

The SRHS question for the MEPS reads: ``In general, compared to other people of your age, would you say that your health is excellent, very good, good, fair, or poor?''  Of the three surveys used for this paper, the MEPS is the only one to qualify SRHS with a reference to the individual's age, possibly yielding different results from the PSID and HRS and reducing comparability.  In addition to the standard five waves, the MEPS also mails a short self-administered questionnaire (SAQ) to all respondents a few weeks after their second and fourth wave interviews.  The MEPS-SAQ includes a SRHS question that is worded nearly identically to that of the PSID and HRS; see Section \ref{sec:Mixed} below.

Although the latent health model can be applied to minors, the motivation of the paper is to provide a more accurate health process for dynamic structural models, which usually have agents that begin no earlier than age 18.  I thus set the age of model entry (denoted $\AgeMin$ in Section \ref{sec:Model}) at 18, ignoring any observations at younger ages.\footnote{MEPS respondents thus might ``age into'' the sample in the middle of their panel.}  The MEPS censors respondents age at 85, so I drop all individuals who report this in their first wave, as their true age cannot be known.  To match the structure of the survey, the age increment for the model is $\AgeIncr=0.5$ years.  Each respondent's age in their first observed wave is incremented by half a year if they report being older in the next wave, six months later.

Even with its short duration, the MEPS dataset is by far the largest constructed for this paper, both by number of individuals (about 236,000) and total observed waves (1.2 million).  Just under 90\% of MEPS respondents are observed in all five waves of their panel; the majority of the remainder are observed for exactly three periods (one year), possibly representing planned attrition by the survey designers.  However, because the vast majority of respondents are non-elderly and relatively healthy and are observed for only two years, fewer than 2,500 deaths are observed for men and women combined; standard errors for mortality parameters should be expected to be somewhat larger as a result.